Lesson 3
========================================================

***

### What to Do First?
Notes: Understand the variable most central to your analysis.

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
setwd('C:/Users/saxenar/Downloads/EDA_Course_Materials/EDA_Course_Materials/lesson3')
getwd()
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
install.packages('ggthemes', dependencies = TRUE) 
library(ggthemes)
theme_set(theme_minimal(24)) 
qplot(x=dob_day, data=pf) +
  scale_x_discrete(breaks=1:31)
```

***

#### What are some things that you notice about this histogram?
Response:
People with dob_day = 1 are most and that with 31 are least.
Average falls around 3500 for all days.
***

### Moira's Investigation
Notes:
whether people's perception of their audience on Facebook matches up to the reality.
There was a pretty big diffrence between perception and reality.
***

### Estimating Your Audience Size
Notes:
People under estimated the size of their audience.
***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes: takes a formula inside of its parenthesis. the formula contains a tilde sign followed by a variable you want to facet over.

facet_grid(vertical~horizontal) useful for 2 variables problem

```{r Faceting}
qplot(x=dob_day, data=pf) +
  scale_x_discrete(breaks=1:31)+
  facet_wrap(~dob_month, ncol=3)

```

#### Let’s take another look at our plot. What stands out to you here?
Response:
The number of people with bday on jan 1 is very high which maybe because of default fb settings.
***

### Be Skeptical - Outliers and Anomalies
Notes:
There are many types of anomalies and outliers, a outlier may be accurate data about extreme case.  sometimes anomalies and outliers represent bad data.
***

### Moira's Outlier
Notes:
someone who just randomly predicted 1 million audience
#### Which case do you think applies to Moira’s outlier?
Response:
bad data about an extreme case
***

### Friend Count
Notes: may need to adjust our axes to get a better look at the data.

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x=friend_count, data=pf)
```

#### How is this plot similar to Moira's first plot?
Response:
The data  is squished to left hand side of graph. this is wahat we call long tail data.  
***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x=friend_count, data=pf) +
  scale_x_continuous(limits=c(0,1000))
```

### Exploring with Bin Width
Notes:
bin width help us get a better insight into the data and do a better  analysis
***

### Adjusting the Bin Width
Notes:
```{r}
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data=pf, xlim=c(0,1000),xlab=seq(0,1000,50),binwidth=25)+
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
ggplot(aes(x = friend_count), data = subset(pf, !is.na(gender))) +
  geom_histogram() + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response: female have more friends than males

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: more robust method

***

### Tenure
Notes:

```{r Tenure}
qplot(x=tenure, data=pf, binwidth=30, color=I('black'), fill = I('#099DD9'))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x=tenure/365, data=pf, binwidth=1, color=I('black'), fill = I('#F79420'))

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(x=tenure/365, data=pf, xlab="Number of years on facebook", ylab="number of user in a sample", binwidth=1, color=I('black'), fill = I('#F79420'))+
  scale_x_continuous(breaks=seq(1,7,1),lim=c(0,7))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(x=age, data=pf, binwidth=1, color=I('black'), fill = I('#F79420'))+
  scale_x_discrete(breaks=seq(0,113,5))
```

#### What do you notice?
Response: that the spike is there at 20 and starts from 13 as it is the least age to join facebook. and at the right we have some bogus values.

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes: meme is adapted on facebook. meme is staying on facebook but in smalller numbers and then grows exponentially.

***

### Transforming Data
Notes:
data is over-dispersed. "Over-dispersed" is always relative to some particular posited distribution. For example, data might be over-dispersed compared with a Poisson distribution with that mean. 
***
```{r}
qplot(x=friend_count, data=pf)

summary(pf$friend_count)

summary(log10(pf$friend_count+1))

summary(sqrt(pf$friend_count))
```

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
install.packages('gridExtra')
library(gridExtra) 

p1 <- qplot(x=friend_count, data=pf)
p2 <- qplot(x=log10(friend_count+1), data=pf)
p3 <- qplot(x=sqrt(friend_count), data=pf)

grid.arrange(p1,p2,p3, ncol=1)
```
***


### Frequency Polygons

```{r Frequency Polygons}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) + 
  xlab('Friend Count') + 
  ylab('Percentage of users with that friend count')
```

***

### Likes on the Web
Notes: males have more likes than females. Since the graph produces is tailed one we are going to use log scaled graphs.

After using scaled log10 graph, it shows that females eventually overtake in number of likes count. so we can't be sure who has more likes

```{r Likes on the Web}
qplot(x=www_likes, data=subset(pf, !is.na(gender)), geom='freqpoly', color=gender)+
  scale_x_continuous()+
  scale_x_log10()

by(pf$www_likes,pf$gender,sum)
```


***

### Box Plots
Notes: different point about box plot is that the y axis is going to be out front count while   the x -axis is categorical variable

```{r Box Plots}
qplot(x=gender, y=friend_count, data=subset(pf, !is.na(gender)),geom='boxplot')
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x=gender, y=friend_count, data=subset(pf, !is.na(gender)),geom='boxplot', ylim=c(0,1000))

qplot(x=gender, y=friend_count, data=subset(pf, !is.na(gender)),geom='boxplot')+
  scale_y_continuous(limits=c(0,1000))

qplot(x=gender, y=friend_count, data=subset(pf, !is.na(gender)),geom='boxplot')+
  coord_cartesian(ylim=c(0,250))

by(pf$friend_count, pf$gender, summary)
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(x=gender, y=friendships_initiated,data=subset(pf, !is.na(gender)),geom='boxplot')+
  coord_cartesian(ylim=c(0,520))
```
```

#### On average, who initiated more friendships in our sample: men or women?
Response: women initiated more friendships than men
#### Write about some ways that you can verify your answer.
Response: the box plot shows the women median line to be higher than that of men
```{r Friend Requests by Gender}
by(pf$friendships_initiated, pf$gender, summary)
```

Response: women initiate more friendship than men

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)

summary(pf$mobile_likes>0)

mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes>0,1,0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

sum(pf$mobile_check_in==1)/length(pf$mobile_check_in)
```

Response: start developing mobile interest among users as only 65% user use it

***

### Analyzing One Variable
Reflection: it is important to take a close look at individual variable of dataset, to understand the types of values they take on, outliers and anomaly.

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!