Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:
people perceived size match with actual size?
audience size is actually as compared to guess made by people
***

### Scatterplots
Notes:

```{r Scatterplots}
setwd('C:/Users/saxenar/Downloads/EDA_Course_Materials/lesson4')
getwd()
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv',sep='\t')

qplot(x=age, y=friend_count, data=pf)
qplot(age, friend_count, data=pf)
```

***

#### What are some things that you notice right away?
Response: People with age less than 30 have more friends than other people.
People with higher age or maybe fake accounts which belongs to teenagers again have high number of friends

***

### ggplot Syntax
Notes: almost same syntax as qplot() except has aes() wrapper where you pass x and y axis variable.

```{r ggplot Syntax}
ggplot(aes(x=age, y=friend_count),data=pf) + geom_point() + xlim(13,90)

summary(pf$age)
```

***

### Overplotting
Notes:using jitter we can add some noise and get better reflections

```{r Overplotting}
ggplot(aes(x=age, y=friend_count),data=pf) + geom_jitter(alpha=1/20) + xlim(13,90)
```

#### What do you notice in the plot?
Response:
the user with age<25 have large number of friends as compared to others. 
Also user with age close to 70 have many friends.
***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x=age, y=friend_count),data=pf) + geom_point(alpha=1/20) + xlim(13,90) + coord_trans(y='sqrt')
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x=age, y=friend_count),data=pf) + geom_jitter(alpha=1/20) + xlim(13,90) + coord_trans(y='sqrt')
```

#### What do you notice?
More detailed knowledge of y parameter(zoomed in version)
***

### Alpha and Jitter
Notes:
Jitter can add both positive and negative noise to the graph.

```{r Alpha and Jitter}
ggplot(aes(x=age, y=friendships_initiated),data=pf) + geom_jitter(alpha=1/10)
```

***

### Overplotting and Domain Knowledge
Notes:
transformed the axis. all of the points are below a specific line.
people under estimate their audience
***

### Conditional Means
Notes:
it can be helpful to summarize over two variables rather than plotting for every variables.
```{r Conditional Means}
install.packages('dplyr')
library(dplyr)

age_groups <- group_by(pf,age)

pf.fc_by_age <- summarise(age_groups, friend_count_mean= mean(friend_count), friend_count_median = median(friend_count), n=n())

pf.fc_by_age <- arrange(pf.fc_by_age, age)

head(pf.fc_by_age)
```

```{r Alternative}
pf.fc_by_age <- pf %.%
  group_by(age)%.%
  summarise(age_groups, friend_count_mean= mean(friend_count),
            friend_count_median = median(friend_count), n=n())%.%
  arrange(age)

head(pf.fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x=age, y=friend_count_mean),data=pf.fc_by_age) + 
         geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes: useful for quick exploration and for raw data

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age, y=friend_count),data=pf) + geom_point(alpha=1/20, position=position_jitter(h=0), color='orange') + coord_trans(y='sqrt') + geom_line(stat= 'summary', fun.y=mean)+ geom_line(stat='summary', fun.y= quantile, probs=0.1, linetype=2, color='blue')+geom_line(stat='summary', fun.y= quantile, probs=0.9, linetype=2, color='blue')+geom_line(stat='summary', fun.y= quantile, probs=0.5, color='blue') + coord_cartesian(xlim=c(60,120))
```

#### What are some of your observations of the plot?
Response:
There is a increase in numbr of friends at the age of 15 and then it drops after age 18 till age 22 and then becomes constant.
Also there are peaks at age 69 and then at age betwen 100-110 which is beacuse of the fake accounts of teenagers who have large number of friends.
having more than one thousand friends is quite rare.
Friend count for user between 35 and 60 fall below 250
***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes: histogram helps in capturing more summary data as compared to scatter plots. histogram show how much people under estimate their audience size more accurately.

***

### Correlation
Notes: linear relationship between age and friend count using correlation.

```{r Correlation}
cor.test(pf$age, pf$friend_count, method='pearson')

with(pf, cor.test(age,friend_count, method='pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:
-0.027
***

### Correlation on Subsets
Notes: not linear relationship between age and friend_count

```{r Correlation on Subsets}
with(subset(pf, pf$age<=70), cor.test(age, friend_count))
```
as age increases friend count decreases. one variables doesn't cause the other
***

### Correlation Methods
Notes: monotonic realtionship like spearman which are increasing and decreasing. Also keats' tau method
```{r}
with(subset(pf, pf$age<=70), cor.test(age, friend_count, method='spearman'))
```

***

## Create Scatterplots
Notes: www_likes_received v/s likes_received. likes_received should be greater than www_likes_received as some of the likes are coming from mobile or other sources.

```{r}
ggplot(aes(x=www_likes_received, y=likes_received),data=pf) + geom_point()
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
install.packages('ggplot2')
library(ggplot2)
ggplot(aes(x=www_likes_received, y=likes_received),data=pf) + geom_point() + xlim(0, quantile(pf$www_likes_received, 0.95)) + ylim(0, quantile(pf$likes_received, 0.95)) + geom_smooth(method='lm', color='red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

Response: 0.948
These two variables are very strongly correlated and generally two variables are not that closely related.
***

### Moira on Correlation
Notes:
to really quantify how coorelated two variables are. in regression we generally look at independent variables. So using correlation test we can find out which variables to use for regression. 
***

### More Caution with Correlation
Notes:
even correlation coefficients can be deceptive. it is better to plot the two variables and then check.
```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
str(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x=Month, y=Temp),data=Mitchell) + geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0
b. What is the actual correlation of the two variables?
(Round to the thousandths place)
0.057
```{r Noisy Scatterplots}
cor.test(Mitchell$Month,Mitchell$Temp)

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x=Month, y=Temp),data=Mitchell) + geom_point() + scale_x_discrete(breaks=seq(0,203,12))
```

***

### A New Perspective

What do you notice?
Response:
I noticed a sinusoidal kind of graph which is repeated over time.
Proposition of variables and scale of graph also matter

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes: friend_count increases and then decreases. Age after 80 are noise.

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x=age, y=friend_count_mean),data=pf.fc_by_age) + 
         geom_line()

head(pf.fc_by_age,10)
pf.fc_by_age[17:19,]
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- pf$age + (12-pf$dob_month)/12

```

Programming Assignment
```{r Programming Assignment}
library(dplyr)
pf.fc_by_age_months <- pf%.%
  group_by(age_with_months)%.%
  summarise(friend_count_mean=mean(friend_count),
            friend_count_median=median(friend_count),
            n=n())%.%
  arrange(age_with_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
p2 <- ggplot(aes(x=age_with_months, y=friend_count_mean),data=subset(pf.fc_by_age_months, age_with_months<71)) + geom_line()

p1 <- ggplot(aes(x=age_with_months, y=friend_count_mean),data=pf.fc_by_age) + geom_line()

library(gridExtra)
grid.arrange(p1,p2,rcol=1)
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
ggplot(aes(x=age_with_months, y=friend_count_mean),data=subset(pf.fc_by_age_months, age_with_months<71)) + geom_line()
```

***

### Which Plot to Choose?
Notes: you dont have to choose. you can work on multiple versions at a time.

***

### Analyzing Two Variables
Reflection:
scatter plots, conditional means, correlation tests.
in this lesson how to work with two variables using scatter plots. we also learned the benefits of correlation and how it can help to decide two variables to work upon.
we learned how to use jitter
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

